---
title: "Movilens"
author: "Monica Bustamante"
date: "1/28/2021"
output: pdf_document
---


#Load Data
```{r config, }
df <- "MovieLens.RData"
if(!file.exists("MovieLens.RData"))
{
  print(df)  
}
```

#Prepare dataset:
```{r setup, include=FALSE}

##########################################################
# Create edx set, validation set (final hold-out test set)
##########################################################

# Note: this process could take a couple of minutes

if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")

library(tidyverse)
library(caret)
library(data.table)

# MovieLens 10M dataset:
# https://grouplens.org/datasets/movielens/10m/
# http://files.grouplens.org/datasets/movielens/ml-10m.zip

dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- read.table(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                      col.names = c("userId", "movieId", "rating", "timestamp"))
movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
                                           title = as.character(title),
                                           genres = as.character(genres))
movielens <- left_join(ratings, movies, by = "movieId")


movielens <- left_join(ratings, movies, by = "movieId")
nrow(movielens)
n_distinct(movielens$genres)

# 3. Validation set will be 10% of MovieLens data
set.seed(1, sample.kind="Rounding") # if using R 3.5 or earlier, use `set.seed(1)`
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)

#Training sets
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set
validation <- temp %>% 
      semi_join(edx, by = "movieId") %>%
      semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set
removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)

#rm(dl, ratings, movies, test_index, temp, movielens, removed)

#Validation Dataset
validation <- validation %>% select(-rating)
```


##GENERAL QUESTIONS Movielens DataSet:
##1. How many rows and columns are there in the edx dataset?

```{r MovielensCapstone, echo=FALSE}
head(edx)
```

#2. Dimension Dataset
```{r MovielensCapstone, echo=FALSE}
dim(edx)
str(edx)
```

# 3. General information about dataset
```{r MovielensCapstone, echo=FALSE}
summary(edx)
```

# 4.How many rows and columns are there in the edx dataset
```{r MovielensCapstone, echo=FALSE}
paste('The dataset has',nrow(edx),'rows and',ncol(edx),'columns.')
```

# 5.To see more information about dataset
```{r MovielensCapstone, echo=FALSE}

edx %>% summarise(
  uniq_movies = n_distinct(movieId),
  uniq_users = n_distinct(userId),
  uniq_genres = n_distinct(genres))
```

#6. Mean of rating dataset
```{r MovielensCapstone, echo=FALSE}
rating_mean <- mean(edx$rating)
rating_mean
```

#7. How many zeros were given as ratings in the edx dataset?
```{r MovielensCapstone, echo=FALSE}

paste(sum(edx$rating == 0), '.ratings and',
      sum(edx$rating == 3),'ratings with 3')
```
```{r MovielensCapstone, echo=FALSE}

edx %>% filter(rating == 3) %>% tally()
```

#8. How many different movies are in the edx dataset
```{r MovielensCapstone, echo=FALSE}

n_distinct(edx$movieId)
```

```{r }
edx %>% summarize(n_movies = n_distinct(movieId))
```
#9. How many different users are in the edx dataset. n_distinct or lenght
```{r MovielensCapstone, echo=FALSE}
n_distinct(edx$userId)
```

```{r }
edx %>% summarize(n_users = n_distinct(userId))
```

#How many movie ratings are in each of the following genres in the edx dataset?
# str_detect

```{r MovielensCapstone, echo=FALSE}

genres <- c("Drama", "Comedy", "Thriller", "Romance")
genresCounts <- sapply(genres, function(g){
    movies %>% filter(str_detect(genres, g)) %>% tally()
})
          
# separate_rows, much slower!
movies %>% separate_rows(genres, sep = "\\|") %>%
    group_by(genres) %>%
    summarize(count = n()) %>%
    arrange(desc(count))
```

#Rating by category
```{r MovielensCapstone, echo=FALSE}

drama <- data %>% filter(str_detect(genres,"Drama"))
paste('Drama',nrow(drama),'movies')

comedy <- data %>% filter(str_detect(genres,"Comedy"))
paste('Comedy',nrow(comedy),'movies')

thriller <- data %>% filter(str_detect(genres,"Thriller"))
paste('Thriller',nrow(thriller),'movies')

romance <- data %>% filter(str_detect(genres,"Romance"))
paste('Romance',nrow(romance),'movies')

rm(drama, comedy, thriller, romance)
```


#Which movie has the greatest number of ratings?

#Greatest number of ratings. Arrange rows by variables

```{r MovielensCapstone, echo=FALSE}
n_distinct(movies$genres)
```

```{r MovielensCapstone, echo=FALSE}

movies %>% group_by(title) %>% 
        summarise(number = n()) %>% 
        arrange(desc(number))
```


#What are the five most given ratings in order from most to least?
#Sort a variable in descending order.
```{r MovielensCapstone, echo=FALSE}
edx %>% group_by(rating) %>% 
    summarize(count = n()) %>% 
    top_n(5) %>%
	  arrange(desc(count))
```

```{r MovielensCapstone, echo=FALSE}
head(sort(-table(edx$rating)),5)
```

```{r MovielensCapstone, echo=FALSE}
hist(edx$rating)
summary(edx$rating)
```

#True or False: In general, half star ratings are less common than whole star ratings (e.g., there are fewer ratings of 3.5 than there are ratings of 3 or 4, etc.).
#Rating movies

```{r MovielensCapstone, echo=FALSE}
rating4 <- table(edx$rating)["4"]
rating35 <- table(edx$rating)["3.5"]
rating3 <- table(edx$rating)["3"]

Result <- (rating35 < rating3 && rating35 < rating4)

print(Result)

rm(rating3, rating35,  rating4, Result)
```


#Graphic Rating movies
```{r MovielensCapstone, echo=FALSE}
edx %>%
	group_by(rating) %>%
	summarize(count = n()) %>%
	ggplot(aes(x = rating, y = count)) +
	geom_line()
```


#4 MODELING
##Predicted movie ratings and calculates RMSE.
#Compared the ratings -> using RMSE

```{r MovielensCapstone,  echo=FALSE}
data <- movies %>% separate_rows(genres, sep ="\\|")
DAT.aggregate <- aggregate(formula = cbind(n = 1:nrow(data)) ~ genres, data = data, FUN = length)
```


#Size of dataset 
```{r MovielensCapstone, echo=FALSE}
movielens <- left_join(ratings, movies, by = "movieId")
nrow(movielens)
```

```{r MovielensCapstone, echo=FALSE}
#Creates Year column.
edx <- edx %>% 
  mutate(title = str_trim(title)) %>%
  extract(title, c("title_tmp", "year"),
          regex = "^(.*) \\(([0-9 \\-]*)\\)$",
          remove = F) %>%
  mutate(year = if_else(str_length(year) > 4,
                        as.integer(str_split(year, "-",
                                             simplify = T)[1]),
                        as.integer(year))) %>%
  mutate(title = if_else(is.na(title_tmp), title, title_tmp)) %>%
  select(-title_tmp)  %>%
  mutate(genres = if_else(genres == "(No Genres Listed)",
                          `is.na<-`(genres), genres))
validation <- temp %>% 
  semi_join(edx, by = "movieId") %>%
  semi_join(edx, by = "userId")
```

 
```{r MovielensCapstone, echo=FALSE}
#Root Mean Square Error  
RMSE <- function(true_ratings, predicted_ratings){
        sqrt(mean((true_ratings - predicted_ratings)^2))
      }
      
lambdas <- seq(0, 5, 0.25)
rmses <- sapply(lambdas,function(l){
  mu <- mean(edx$rating) #The mean of ratings from training set
  
  Movie_effect <- edx %>%  #Adjust mean by movie effect
    group_by(movieId) %>%
    summarize(Movie_effect = sum(rating - mu)/(n()+l))
  
  Movie_user <- edx %>% #Ajdust mean by movie effect and user  
    left_join(Movie_effect, by="movieId") %>%
    group_by(userId) %>%
    summarize(Movie_user = sum(rating - Movie_effect - mu)/(n()+l))
  
  predicted_ratings <- 
    edx %>% 
    left_join(Movie_user, by = "userId") %>%
    left_join(Movie_effect, by = "movieId") %>%
    mutate(pred = mu + Movie_effect + Movie_user) %>%
    .$pred #Predict ratings 
  
  return(RMSE(predicted_ratings, edx$rating))
})
plot(lambdas, rmses, col = "blue")
```
#Calculate Lambda optimal RMSE
```{r MovielensCapstone, echo=FALSE}
lambda <- lambdas[which.min(rmses)]
paste('RMSE',min(rmses),'Lambda',lambda)

```

```{r MovielensCapstone, echo=FALSE}
mu_hat <- mean(train_set$rating)
naive_rmse <- RMSE(test_set$rating, mu_hat)
rmse_results <- data_frame(method = "Average", RMSE = naive_rmse)
rmse_results

```
#Graphic Histogram summarize 
```{r MovielensCapstone, echo=FALSE}

#UserId Bias
mu <- mu_hat
movie_avgs <- train_set %>% 
  group_by(movieId) %>% 
  summarize(b_i = mean(rating - mu))
movie_avgs %>% qplot(b_i, geom ="histogram", bins = 10, data = ., color = I("blue"))

```

```{r MovielensCapstone, echo=FALSE}

# prediction 
b_i <- test_set %>% 
  left_join(movie_avgs, by='movieId') %>%
  .$b_i

#Delete NA
b_i[is.na(b_i)] <- 0
predicted_ratings <- mu + b_i
model_1_rmse <- RMSE(predicted_ratings, test_set$rating)
rmse_results <- bind_rows(rmse_results,
                          data_frame(method="Movie Model",
                                     RMSE = model_1_rmse ))
rmse_results
```

```{r MovielensCapstone, echo=FALSE}

user_avgs <- train_set %>% 
  left_join(movie_avgs, by='movieId') %>%
  group_by(userId) %>%
  filter(n()>100) %>%
  summarize(b_u = mean(rating - mu - b_i))

#Delete NA
user_avgs$b_u[is.na(user_avgs$b_u)] <- 0
user_avgs %>% qplot(b_u, geom ="histogram", bins = 10, data = ., color = I("green"))
```

```{r MovielensCapstone, echo=FALSE}

b_i <- test_set %>% 
  left_join(movie_avgs, by='movieId') %>%
  .$b_i

#Delete NA
b_i[is.na(b_i)] <- 0
predicted_ratings <- mu + b_i
model_rmse <- RMSE(predicted_ratings, test_set$rating)
rmse_results <- bind_rows(rmse_results,
                          data_frame(method="Movie Model",
                                     RMSE = model_rmse ))
rmse_results
```


#CONCLUSION:

Predict a list of rated movies.

Discovered patterns: as people prefer movies with a medium to high rating. (3 to 5). 

The movies preferred by the customers was the end of the 1980 and 1990 periods.

